Thesis: draft

### 2021-04-09 draft.tex
latest update

The initial fraction of nodes with opinion 0 is now equal to $P_0(0) = 0.2$, opinion 0 thus becomes the minority opinion. The initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.8$, hence this is the majority opinion. The opinions are, initially, randomly distributed over the system. Again, the opinion dynamics model of Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented; the activation probability is set equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length $S = 20$.\\
\newline
Once again, we look at four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $10$
\begin{itemize}
	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$. The value of the average clustering coefficient is equal to $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
	\newpage
	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.
	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100. Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$. Edges are added between communities with a probability $p_{\text{add}} = 0.001$. This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
\end{itemize}
\textbf{Evolution of the opinion fractions}\\
\newline
Figure \ref{ev_op_20_80} shows the results for the evolution of the fraction of nodes with opinion 0, which is the minority opinion in this case, over time, together with the standard deviation $\sigma$. For each network model two filtering algorithms are compared, namely the REC filtering algorithm and the PR filtering algorithm. \\
The REC method is not able to break the prevalence of opinion 0 over time for neither of the four different network models. When using the PR method, the results look very different. The prevalence of the minority opinion, $P_0(t)$, decreases over time. This decrease is a direct consequence of the PR filtering algorithm. Since the PR method orders the posts according to the users' preferences, nodes with the majority opinion 1 will mainly get to see posts with the similar opinion 1 on their time-line, which limits the chance of changing their opinion to the minority opinion. Nodes with the minority opinion 0, on the other hand, will obviously see posts with opinion 0 on their time-line when using the PR method. However, since there are more nodes with opinion 1 than nodes with opinion 0, the nodes with the minority opinion 0 will also get to see a significant number of posts with the majority opinion 1 on their time-line. The conclusion is that, when using the PR method, there is an imbalance in which opinion can switch to which: nodes with the minority opinion 0 have a chance of switching to the majority opinion 1, whereas nodes with the majority opinion 1 have (almost) no chance of switching to the minority opinion 0. This imbalance leads to the decrease of the minority opinion 0.

@@

The PR filtering algorithm is compared to the REC filtering algorithm. For most values of the modularity, the REC method produces almost no echo chambers. If the modularity becomes high enough we start to see a very small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than 2, which means that after 500 time steps there are less than two times as much echo chambers than in the beginning at $t=0$.\\

@@

\subsection{Including a stubbornness parameter}
\label{stubbpar}
Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
\begin{itemize}
	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
\end{itemize}
For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around 10)
\begin{itemize}
	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$, with $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.
	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100. Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$. Edges are added between communities with a probability $p_{\text{add}} = 0.001$. This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
\end{itemize}
For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.\\
\newline
\textbf{Making a fraction of the nodes completely stubborn}\\
\newline
nsdjk\\
\newline
\textbf{Making all nodes partly stubborn}\\
\newline
jsqnbd\\

>>>>

The initial fraction of nodes with opinion 0 is now equal to $P_0(0) = 0.2$, opinion 0 thus becomes the minority opinion. The initial fraction of nodes with opinion 1 is equal to $P_1(0) = 0.8$, hence this is the majority opinion. The opinions are, initially, randomly distributed over the system. Again, the opinion dynamics model of Chapter \ref{chap4}, Section \ref{opinionmodel} is implemented; the activation probability is set equal to $p_{\text{act}} = 0.1$ and the actual time-line of each user has a maximum length $S = 20$.\\
Once again, we look at four different network models each with a total number of nodes equal to $N = 10^3$ and an average degree around $10$
	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$. The value of the average clustering coefficient is equal to $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
Figure \ref{ev_op_20_80} shows the results for the evolution of the fraction of nodes with opinion 0, which is the minority opinion in this case, over time, together with the standard deviation $\sigma$. For each network model two filtering algorithms are compared, namely the REC filtering algorithm and the PR filtering algorithm. \\
The REC method is not able to break the prevalence of opinion 0 over time for neither of the four different network models. When using the PR method, the results look very different. The prevalence of the minority opinion, $P_0(t)$, decreases over time. This decrease is a direct consequence of the PR filtering algorithm. Since the PR method orders the posts according to the users' preferences, nodes with the majority opinion 1 will mainly get to see posts with the similar opinion 1 on their time-line, which limits the chance of changing their opinion to the minority opinion. Nodes with the minority opinion 0, on the other hand, will obviously see posts with opinion 0 on their time-line when using the PR method. However, since there are more nodes with opinion 1 than nodes with opinion 0, the nodes with the minority opinion 0 will also get to see a significant number of posts with the majority opinion 1 on their time-line. The conclusion is that, when using the PR method, there is an imbalance in which opinion can switch to which: nodes with the minority opinion 0 have a chance of switching to the majority opinion 1, whereas nodes with the majority opinion 1 have (almost) no chance of switching to the minority opinion 0. This imbalance leads to the decrease of the minority opinion 0.\\
From Figure \ref{ev_op_20_80}, it is clear that the minority opinion 0 eventually ceases to exist for all the network models, $P_0(500) \sim 0$. This result differs from the conclusion that prof. N. Perra and prof. L.E.C. Rocha made in the paper \textit{"Modelling opinion dynamics in the age of algorithmic personalization"} (see reference \cite{Perra2019}). They found that the minority opinion was able to survive over time, even for the PR method, and reported a minority opinion fraction at $t=500$ between $P_0(500) = 0.06$ and $P_0(500) = 0.09$ \cite{Perra2019}. % explain further, but first test WS, K=6, beta=0.01, N=1000, 20-80, PR

@@

The PR filtering algorithm is compared to the REC filtering algorithm. For most values of the modularity, the REC method produces almost no echo chambers. If the modularity becomes high enough we start to see a very small increase in the formation of echo chambers, however even for $Q \sim 0.9$, the size of the echo chambers remains smaller than 2, which means that after 500 time steps there are less than twice as much echo chambers than in the beginning at $t=0$.\\

@@

\subsection{Including a stubbornness parameter}
\label{stubbpar}
Here, we review the results of introducing stubborn people by including a stubbornness parameter $r_i$. We look at two slightly different case
\begin{itemize}
	\item Making a fraction of the nodes (going from none of the nodes to all the nodes) completely stubborn ($r_i = 1$),
	\item Making all nodes partly stubborn (going from not stubborn at all, $r_i = 0$ to completely stubborn, $r_i = 1$).
\end{itemize}
For both cases we will look at both the evolution of the opinions and the formation of echo chambers. \\
We will again implement four different network models in order to investigate the possible interplay between network structure and stubbornness, these models are (each with a total number of nodes $N = 10^3$ and an average degree around 10)
\begin{itemize}
	\item The Erd\H{o}s-R\'{e}nyi model (ER) with edge probability $p = 0.01$.
	\item The Watts-Strogatz model (WS) with mean degree $K=10$ and rewire probability $\beta = 0.06$, with $\left<C\right> \sim 0.55$. This model is used to investigate the effect of clustering.
	\item The stochastic block model with 10 communities of size 100; two networks with different modularities $Q$ are investigated: a high modularity network ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity network ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$). This model is used to investigate the effect of a community structure.
	\item The stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100. Each community has the following Watts-Strogatz parameters: mean degree $K =10$ and rewire probability $\beta = 0.01$. Edges are added between communities with a probability $p_{\text{add}} = 0.001$. This is a network with both a high clustering ($\left<C\right> \sim 0.55$) and a high modularity ($Q \sim 0.9$).
\end{itemize}
For each network model, we compare the REF, REC and PR method to each other to analyze the possible interplay between stubbornness and filtering algorithm.\\
\newline
\textbf{Making a fraction of the nodes completely stubborn}\\
\newline
We start by making a fraction of the nodes (fracRes) completely stubborn ($r_i = 1$). We go from no stubborn nodes, fracRes $=0$, to all the nodes are stubborn, fracRes $=1$.\\
\newline
\textbf{Making all nodes partly stubborn}\\
\newline
jsqnbd\\
